General constraints on magic distillation protocols from statistical mechanics

### 2021-06-06 magic_paper.tex
Restructiing of appendices.

We note that we may relate our analysis to the classical single-shot entropy $D_\infty (\p ||\r)$ that remains well-defined even in the case of quasi-distributions. We define 
\begin{equation}
	\hspace{-14pt}D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x)) = \log  \max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)},
\end{equation}
which is well-defined since both the nominator and denominator in the above expression are strictly positive when $\tau$ is in the interior of $\F$.

\newpage

\subsection{Entropic bounds on quasi-distributions}

We prove the following result connecting simple majorization to relative majorization.
\nick{move following lemma to section on relative majorization equivalent conditions}
\begin{lemma}
Let $\w\in \mathbb{R}^N, \w'\in \mathbb{R}^{N'}$ be quasi-distributions and $\r\in \mathbb{R}^N, \r'\in \mathbb{R}^{N'}$ probability distributions with positive rational entries given by $r_i = a_i/K$ and $r_i' = a_i'/K$ for positive integers $a_i, a_i'$ and $K = \sum_{i=1}^N a_i = \sum_{i=1}^{N'} a_i'$. 
Then,
\begin{equation}
(\w, \r) \succ (\w', \r') \mbox{ if and only if } \Gamma_{\bma} (\w) \succ \Gamma_{\bma'} (\w'),
\end{equation}
where the embedding map $\Gamma_{\bma} :\mathbb{R}^N \rightarrow \mathbb{R}^K$ is given by
\begin{equation}
	\Gamma_{\bma}(\x) := \bigoplus_{i=1}^N x_i \eta_{a_i},
\end{equation}
with $\eta_{a_i} = (1/a_i, 1/a_i, \dots, 1/a_i)$ the uniform distribution on $a_i$ elements.
\end{lemma}
\begin{proof}
	We first note that the map $\Gamma_{\bma}$ is stochastic and has a well-defined left-inverse $\Gamma^{-1}_{\bma'}: \mathbb{R}^K \rightarrow \mathbb{R}^N$ given by
\begin{equation}
	\Gamma^{-1}_{\bma}(\w) = \left( \sum_{i=1}^{a_1} w_i,  \sum_{i = a_1 +1}^{a_1 + a_2} \hspace{-5pt} w_i, \dots , \hspace{-15pt} \sum_{i=a_1 + \dots + a_{N-1}+1}^{K} \hspace{-10pt} w_i \right),
\end{equation}
that obeys $(\Gamma^{-1}_{\bma} \circ \Gamma_{\bma}) (\x) = \x$ for all $\x \in \mathbb{R}^N$.

\subsubsection{Classical $\alpha$-Renyi entropy and divergence for quasi-distributions}

We first define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.
Note that for any quantum state $\rho$, $H_1(\rho)$ in particular is not well-defined  \nick{Notation clashes with quantum Renyi entropies}, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}

We can prove additivity of $H_{\alpha}$ on quantum states.

\begin{lemma}\label{H_add}
	For any quantum states $\rho, \rho'$, we have
	\begin{equation}
		H_{\alpha}(\rho \otimes \rho') = H_{\alpha}(\rho) + H_{\alpha}(\rho'),
	\end{equation}
	where $\hspace{-1pt}\alpha\hspace{-1pt}$ is such that $\hspace{-1pt}H_\alpha(\sigma)\hspace{-2pt}$ is well-defined for all states $\sigma$.
\end{lemma}
\begin{proof}
	This can be seen directly by the multiplicative property of the Wigner distribution,
	\begin{align*}
		H_{\alpha}&(\rho \otimes \rho') = \frac{1}{1-\alpha}\log \sum_{\bmx,\bmx'} \left[W_{\rho}(\bmx)W_{\rho}(\bmx')\right]^{\alpha} \\
		&= \frac{1}{1-\alpha}\left( \log \sum_{\bmx}W_{\rho}(\bmx)^{\alpha} + \log \sum_{\bmx'}W_{\rho}(\bmx')^{\alpha} \right) \\
		&= H_{\alpha}(\rho) + H_{\alpha}(\rho').
	\end{align*}
\end{proof}

We now have the following theorem~\cite{cit:marshall} that provides a characterization of symmetric, Schur-convex and Schur-concave functions on quasi-distributions.
\begin{lemma}
Let $f$ be a real-valued continuous function defined on $\mathcal{Q}^\downarrow$ that is continuously differentiable on the interior of $\mathcal{Q}^\downarrow$. 
Then, $f$ is Schur-convex on $\mathcal{Q}^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \mathcal{Q}^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\mathcal{Q}^\downarrow$.
\end{lemma}

Now we define the classical $\alpha$-Renyi divergence in terms of the Wigner quasi-distribution of any magic state $\rho$ and the Wigner probability distribution of any equilibrium state $\tau$ as
\begin{equation}
	D_\alpha(\rho || \tau) := \frac{1}{\alpha - 1} \log_2 \sum_{\bmx} W_{\rho}(\bmx)^\alpha W_{\tau}(\bmx)^{1-\alpha},
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.
\nick{Notation clashes with quantum Renyi divergence}

\begin{theorem}\label{thm:DSchur}
	Let $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$. 
	Then, $D_\alpha$ is Schur-convex, i.e. let $\w, \w' \in \mathbb{R}^N$ be quasi-distributions and $\r, \r' \in \mathbb{R}^N$ be probability distributions with positive rational entries given by $r_i = a_i/K$ and $r_i' = a_i'/K$ for positive integers $a_i, a_i'$ and $K = \sum_{i=1}^N a_i = \sum_{i=1}^{N'} a_i'$.
	Then, we have $D_\alpha(\w || \r) \geq H_\alpha(\w' || \r')$, whenever $(\w, \r) \succ (\w', \r')$.
\end{theorem}
\begin{proof}
	The statement $(\w, \r) \succ (\w', \r')$ is equivalent to $\Gamma_{\bma}(\w) \succ \Gamma_{\bma'}(\w')$.
	Therefore, due to~\cref{thm:HSchur} we have
	\begin{equation}
		H_\alpha(\Gamma_{\bma}(\w)) \leq H_\alpha(\Gamma_{\bma'}(\w'))
	\end{equation}
	which due to~\cref{thm:H2D} is equivalent to
	\begin{equation}
		D_\alpha(\w ||\r) \geq D_\alpha(\w' ||\r').
	\end{equation}
\end{proof}

\ddd{Therefore if we have a Schur-Concave function $f(\x)$ for majorization of quasi-distributions then we can define a monotonic function for relative majorization via composition with $\Gamma_{\mathbf{a}}$ and its inverse.
This should provide relative majorization monotones from which we can obtain distillation bounds, such as
\begin{equation}
	R \le \frac{F_\alpha(\tau) - F_\alpha (\rho)}{F_\alpha(\tau') - F_\alpha (\rho')}.
\end{equation}}
\nick{I don't understand where the free energies came from. Are they known to be Schur-concave?}

	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{D_{\alpha}(\rho||\tau)}{D_{\alpha}(\rho'||\tau')},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$.
\end{theorem}
\begin{proof}
	For the sake of clarity, we write $\rho_n \coloneqq \rho_S(\epsilon)^{\otimes n}$, $\rho'_{n'} \coloneqq \rho_S(\epsilon')^{\otimes n'}$, $\tau_n \coloneqq \tau^{\otimes n}$ and $\tau'_{n'} \coloneqq \tau'^{\otimes n'}$.

\ddd{The following is a nice thing to establish.}
\begin{theorem} A state $\rho$ has Wigner negativity if and only if $H_\alpha (W_\rho (\x)) < 0$ for some $\alpha >1$.
\end{theorem}
Let's try to prove this.
\begin{proof} If $H_\alpha (W_\rho(\x) <0 $ then this implies that $\log \sum_{\x} W_\rho (\x)^\alpha > 1$, however it is known that $H_\alpha$ is always non-negative on probability distributions and therefore $W_\rho(\x)$ must be a strict quasi-distribution with negativity.

Conversely, suppose $\rho$ has negativity in its Wigner representation. This implies that $\sum_{\x} |W_\rho (\x)| >1$. We now consider $\sum_{\x} W_\rho (\x)^{a/b}$ for $a/b >1$ and $a$ even and $b$ odd. We have that
\begin{align}
\sum_{\x} W_\rho (\x)^{a/b} &= \sum_{\x} |W_\rho (\x)|^{a/b} \nonumber \\
&= \sum_{\x} |W_\rho (\x)|^{1+ \epsilon},
\end{align}
where $\epsilon = a/b - 1 >0$. By choosing $a$ and $b$ appropriately we can make $\epsilon$ arbitrarily close to $0$ and therefore there exists a sequence $\epsilon_n$ such that
\begin{equation}
\sum_{\x} |W_\rho (\x)|^{1+ \epsilon_n} \rightarrow \sum_{\x} |W_\rho (\x)| >1.
\end{equation}
Therefore, $H_\alpha (W_\rho (\x)) <0 $ at some point in the sequence, which completes the proof.
\end{proof}
Therefore Wigner negativity coincides with negativity in at least one Renyi entropy.
\ddd{[Looking at this limit it almost feels like the $\alpha \rightarrow 0^+$ limit is actually giving mana....modulo the technical problem of taking the limit without calculus. I will think more about this point.]}

\subsubsection{Extention to general relative majorization}

Therefore if we have a Schur-Concave function $f(\x)$ for majorization of quasi-distributions then we can define a monotonic function for relative majorization via composition with $\Gamma_{\mathbf{a}}$ and its inverse.

@@

>>>>

@@

\newpage
\section{Single-shot entropies on quasi-distributions and magic distillation bounds}

We note that we may relate our analysis to the classical single-shot entropies that remain well-defined even in the case of quasi-distributions.

\subsection{Renyi min-divergence from initial slope}
We first demonstrate that the initial slope analysis on the Wigner quasi-distributions makes use of the min-relative divergence. We define 
\begin{equation}
	\hspace{-14pt}D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x)) = \log  \max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)},
\end{equation}
which is well-defined since both the nominator and denominator in the above expression are strictly positive when $\tau$ is in the interior of $\F$.

\subsection{Well-defined Renyi entropies on quasi-distributions}

\ddd{[ISSUE WITH NOTATION: We can't clash with the quantum Renyi entropies. I think we have to use $H_\alpha(W_\rho)$ is bulky...but hopefully clear.]}

We define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(W_\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.
Note that for any quantum state $\rho$, $H_1(W_\rho)$ in particular is not well-defined, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(W_\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}

We now have that the Schur-Ostrowski theorem~\cite{cit:marshall} provides a characterization of symmetric, Schur-convex and Schur-concave functions on quasi-distributions.
\begin{lemma}
Let $f$ be a real-valued continuous function defined on $\mathcal{Q}^\downarrow$ that is continuously differentiable on the interior of $\mathcal{Q}^\downarrow$. 
Then, $f$ is Schur-convex on $\mathcal{Q}^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \mathcal{Q}^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\mathcal{Q}^\downarrow$.
\end{lemma}

Now we define the classical $\alpha$-Renyi divergence in terms of the Wigner quasi-distribution of any magic state $\rho$ and the Wigner probability distribution of any equilibrium state $\tau$ as
\begin{equation}
	D_\alpha(\rho || \tau) := \frac{1}{\alpha - 1} \log_2 \sum_{\bmx} W_{\rho}(\bmx)^\alpha W_{\tau}(\bmx)^{1-\alpha},
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.

We can check additivity of $H_{\alpha}$ also holds on quasi-distributions. 
\begin{lemma}\label{H_add}
	For any $\w, \w' \in \mathcal{Q}$, we have
	\begin{equation}
		H_{\alpha}(\w\otimes \w') = H_{\alpha}(\w) + H_{\alpha}(\w'),
	\end{equation}
	where $\alpha = 2a/(2b-1)$ with integers $a\ge b$.
\end{lemma}
\begin{proof}
	That fact that negativity does not affect additivity can be seen directly as follows
	\begin{align*}
		H_{\alpha}(\w \otimes \w') &= \frac{1}{1-\alpha}\log \sum_{i,j'} \left[\w_i \w'_j\right]^{\alpha} \\
		&= \frac{1}{1-\alpha}\log  \left [ \sum_i\w_i^\alpha \sum_j\w_j^{\prime \alpha} \right ] \\
				&= \frac{1}{1-\alpha}\log  \left [ \sum_i\w_i^\alpha\right ] + \frac{1}{1-\alpha}\log \left [ \sum_j\w_j^{\prime \alpha} \right ] \\
		&= H_{\alpha}(\w) + H_{\alpha}(\w').
	\end{align*}
\end{proof}

\subsection{Renyi divergences on quasi-distributions}
We prove the following result connecting majorization to relative majorization.
\begin{lemma}
Let $\w\in \mathbb{R}^N, \w'\in \mathbb{R}^{N'}$ be quasi-distributions and $\r\in \mathbb{R}^N, \r'\in \mathbb{R}^{N'}$ probability distributions with positive rational entries given by $r_i = a_i/K$ and $r_i' = a_i'/K$ for positive integers $a_i, a_i'$ and $K = \sum_{i=1}^N a_i = \sum_{i=1}^{N'} a_i'$. 
Then,
\begin{equation}
(\w, \r) \succ (\w', \r') \mbox{ if and only if } \Gamma_{\bma} (\w) \succ \Gamma_{\bma'} (\w'),
\end{equation}
where the embedding map $\Gamma_{\bma} :\mathbb{R}^N \rightarrow \mathbb{R}^K$ is given by
\begin{equation}
	\Gamma_{\bma}(\x) := \bigoplus_{i=1}^N x_i \eta_{a_i},
\end{equation}
with $\eta_{a_i} = (1/a_i, 1/a_i, \dots, 1/a_i)$ the uniform distribution on $a_i$ elements.
\end{lemma}
\begin{proof}
	We first note that the map $\Gamma_{\bma}$ is stochastic and has a well-defined left-inverse $\Gamma^{-1}_{\bma'}: \mathbb{R}^K \rightarrow \mathbb{R}^N$ given by
\begin{equation}
	\Gamma^{-1}_{\bma}(\w) = \left( \sum_{i=1}^{a_1} w_i,  \sum_{i = a_1 +1}^{a_1 + a_2} \hspace{-5pt} w_i, \dots , \hspace{-15pt} \sum_{i=a_1 + \dots + a_{N-1}+1}^{K} \hspace{-10pt} w_i \right),
\end{equation}
that obeys $(\Gamma^{-1}_{\bma} \circ \Gamma_{\bma}) (\x) = \x$ for all $\x \in \mathbb{R}^N$.

We now establish monotonicity for Renyi relative divergences.
\begin{theorem}\label{thm:DSchur}
	Let $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$. 
	Then, $D_\alpha$ is Schur-convex, i.e. let $\w, \w' \in \mathbb{R}^N$ be quasi-distributions and $\r, \r' \in \mathbb{R}^N$ be probability distributions with positive rational entries given by $r_i = a_i/K$ and $r_i' = a_i'/K$ for positive integers $a_i, a_i'$ and $K = \sum_{i=1}^N a_i = \sum_{i=1}^{N'} a_i'$.
	Then, we have $D_\alpha(\w || \r) \geq H_\alpha(\w' || \r')$, whenever $(\w, \r) \succ (\w', \r')$.
\end{theorem}
\begin{proof}
	The statement $(\w, \r) \succ (\w', \r')$ is equivalent to $\Gamma_{\bma}(\w) \succ \Gamma_{\bma'}(\w')$.
	Therefore, due to~\cref{thm:HSchur} we have
	\begin{equation}
		H_\alpha(\Gamma_{\bma}(\w)) \leq H_\alpha(\Gamma_{\bma'}(\w'))
	\end{equation}
	which due to~\cref{thm:H2D} is equivalent to
	\begin{equation}
		D_\alpha(\w ||\r) \geq D_\alpha(\w' ||\r').
	\end{equation}
\end{proof}

	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{D_{\alpha}(\rho||\tau)}{D_{\alpha}(\rho'||\tau')} = \frac{F_\alpha(\rho) - F_\alpha( \tau)}{F'_\alpha(\rho') - F'_\alpha( \tau')},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and $\beta F_\alpha(\rho) := D_\alpha (\rho || \tau) - \log \Z$ is the $\alpha$--free energy. 
\end{theorem}
\begin{proof}
	For the sake of clarity, we write $\rho_n \coloneqq \rho_S(\epsilon)^{\otimes n}$, $\rho'_{n'} \coloneqq \rho_S(\epsilon')^{\otimes n'}$, $\tau_n \coloneqq \tau^{\otimes n}$ and $\tau'_{n'} \coloneqq \tau'^{\otimes n'}$.

We now show the following simple equivalence.
\begin{theorem} A state $\rho$ has Wigner negativity if and only if $H_\alpha (W_\rho (\x)) < 0$ for some $\alpha >1$.
\end{theorem}
Let's try to prove this.
\begin{proof} If we have $H_\alpha (W_\rho(\x)) <0 $ then this implies that $\log \sum_{\x} W_\rho (\x)^\alpha > 1$, however it is known that $H_\alpha$ is always non-negative on probability distributions and therefore $W_\rho(\x)$ must be a strict quasi-distribution with negativity.

Conversely, suppose $\rho$ has negativity in its Wigner representation. This implies that $\sum_{\x} |W_\rho (\x)| >1$. We now consider $\sum_{\x} W_\rho (\x)^{a/b}$ for $a/b >1$ and $a$ even and $b$ odd. We have that
\begin{align}
\sum_{\x} W_\rho (\x)^{a/b} &= \sum_{\x} |W_\rho (\x)|^{a/b} \nonumber \\
&= \sum_{\x} |W_\rho (\x)|^{1+ \epsilon},
\end{align}
where $\epsilon = a/b - 1 >0$. By choosing $a$ and $b$ appropriately we can make $\epsilon$ arbitrarily close to $0$ and therefore there exists a sequence $\epsilon_n=a_n/b_n$ such that
\begin{equation}
\sum_{\x} |W_\rho (\x)|^{1+ \epsilon_n} \rightarrow \sum_{\x} |W_\rho (\x)| >1.
\end{equation}
Therefore, $H_\alpha (W_\rho (\x)) <0 $ at some point in the sequence, which completes the proof.
\end{proof}
Therefore Wigner negativity coincides with negativity of a Renyi entropy.

\ddd{[Looking at this limit it almost feels like the $\alpha \rightarrow 1^+$ limit is actually giving mana....modulo the technical problem of taking the limit without calculus. I will think more about this point.]}
